Repeated loss of variation in insect ovary morphology highlights the role of developmental constraint in life-history evolution - Supplementary Methods

### 2020-12-08 manuscript_supplement.Rmd
updating commit number
quick correction in manuscript file
updating Rmd files with new method for calculating and reporting p-values

---
title: Repeated loss of variation in insect ovary morphology highlights the role of developmental constraint in life-history evolution - Supplementary Methods
output: 
  bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: pdflatex
    toc: no
  #bookdown::html_document2: default
bibliography: manuscript_supplement.bib
csl: nature.csl
link-citations: yes
header-includes:
 \usepackage{float}
 \floatplacement{figure}{H}
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(echo=FALSE, include = FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

```{r preliminaries,cache=FALSE}
library(BAMMtools)
library(phytools)
library(coda)
library(OUwie)
library(corHMM)
library(grid)
library(gridExtra)
library(gtable)
library(geiger)
library(arbutus)
library(dplyr)
library(ggtree)
library(ape)
library(ggplot2)
library(RColorBrewer)
theme_set(theme_classic())
options(tinytex.clean = FALSE)
```

@@

```{r load_data, cache=FALSE}
# This script reads in and formats the ovariole number dataset
source("analysis/get_ovariole_number_data.R")
```

Samuel H Church^1^\*, Bruno A. S. De Medeiros^1,2^, Seth Donoughe^1,3^, Nicole L. Márquez Reyes^4^, Cassandra G. Extavour^1,5^\*

@@

Using this approach, we gathered `r nrow(raw_ON)` records for ovariole number from 460 publications. A full list of publications is provided in the supplementary files 'ovariole_number_bibliography.pdf'. We matched the scientific names to additional taxonomic information using the software TaxReformer[@TaxReformer] and found additional taxonomic data for `r nrow(matched)` of the `r nrow(raw_ON)` records. We verified that TaxReformer had found a valid match by comparing the originally recorded taxonomic order to the order populated by online databases, and removed `r nrow(raw_ON %>% filter(problem == "taxonomy"))` taxonomic records for which these values did not match. For all subsequent analyses, we also excluded observations made in non-reproductive individuals from eusocial species (workers), as well as two observations which represented significant outliers and could not be validated using additional sources or figures[@su2015testicular;@hernandez2017biology].

@@

Average adult body length per insect family[@rainford2016phylogenetic] was matched to the average ovariole number for the corresponding families (Fig. \ref{fig:PGLS-egg-ON-family}). The Rainford et al, 2016[@rainford2016phylogenetic] dataset contains a small number of average adult body lengths at the order level (e.g. Strepsiptera), which were matched to their equivalent group in the ovariole number dataset. To test the effect of uncertainty in the estimated average ovariole number on our results, the dataset for each family was downsampled by half at each iteration of the regression analysis.

@@

```{r PGLS-posterior, include=TRUE, cache=FALSE}
#source("analysis/run_pgls_over_posterior_distribution.R")
load("analysis/pgls_results_June2020/run_pgls_over_posterior_distribution.RData")

# analysis names in table 
pgls_analysis_names <- c("ovariole number vs egg volume",
  "ovariole number vs egg volume",
  "ovariole number vs egg volume, residuals to body length",
  "ovariole number vs egg volume, residuals to body length, excluding Trigonalidae",
  "ovariole number vs egg volume, residuals to body weight",
  "ovariole number vs egg volume, residuals to body weight",
  "ovariole number vs lifetime fecundity",
  "ovariole number vs lifetime fecundity",
  "Curculionoidea ovariole number vs egg volume, residual to elytra length",
  "Orthoptera ovariole number vs egg volume, residuals to body length",
  "Hymenoptera ovariole number vs egg volume, residuals to mesosoma width",
  "Drosophilidae ovariole number vs egg volume, residuals to thorax length")
# level at which each PGLS was performed 
taxonomic_levels <- c("species","genus","family","family","species","genus","species","genus","genus","genus","genus","species")

@@

# write table
all_table <- as.data.frame(do.call(rbind, allometry_tables)) %>% round(.,3) %>% mutate(analysis = pgls_analysis_names, "taxonomic level" = taxonomic_levels,slope = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
      intercept = paste(sprintf("%.3f",int_min),sprintf("%.3f",int_max),sep=sp),
      "p-value" = paste(sprintf("%.3f",pval_min),gsub("0.000","<0.000",sprintf("%.3f",pval_max)),sep=sp),
      "taxa" = taxa) 
knitr::kable(all_table  %>% select(c("analysis","taxonomic level","slope","intercept","p-value","taxa")), "latex",caption="Results of PGLS analysis of ovariole number, egg size, and fecundity across a posterior distribution", booktabs = T, linesep = "") %>%
  kableExtra::kable_styling(latex_options = "striped") %>%
  kableExtra::column_spec(1, width = "16em") %>%
  kableExtra::kable_styling(latex_options = "hold_position")  %>%
  kableExtra::kable_styling(font_size = 8) 
```

```{r PGLS-body-posterior, include=TRUE, cache=FALSE}
#source("analysis/run_pgls_body_ovariole_number.R")
load("analysis/pgls_results_June2020/run_pgls_body_ovariole_number.RData")

# analysis names in table 
pgls_analysis_names <- c("ovariole number vs body length",
  "ovariole number vs body weight",
  "ovariole number vs body weight",
  "Curculionoidea ovariole number vs elytra length",
  "Orthoptera ovariole number vs body length",
  "Hymenoptera ovariole number vs mesosoma width",
  "Drosophilidae ovariole number vs thorax length")
# level at which each PGLS was performed 
taxonomic_levels <- c("family","species","genus","genus","genus","genus","species")

# write table 
body_all_table <- as.data.frame(do.call(rbind, body_allometry_tables)) %>% round(.,3) %>% mutate(analysis = pgls_analysis_names, "taxonomic level" = taxonomic_levels,slope = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
       intercept = paste(sprintf("%.3f",int_min),sprintf("%.3f",int_max),sep=sp),
       "p-value" = paste(sprintf("%.3f",pval_min),gsub("0.000","<0.000",sprintf("%.3f",pval_max)),sep=sp),
       "taxa" = taxa)
knitr::kable(body_all_table %>% select(c("analysis","taxonomic level","slope","intercept","p-value","taxa")), "latex",caption="Results of PGLS analysis of ovariole number and body size across a posterior distribution", booktabs = T, linesep = "") %>%
  kableExtra::kable_styling(latex_options = "striped") %>%
  kableExtra::column_spec(1, width = "16em") %>%
  kableExtra::kable_styling(latex_options = "hold_position")  %>%
  kableExtra::kable_styling(font_size = 8) 
```

For each regression, we repeated a Phylogenetic Generalized Least Squares (PGLS) analysis 1000 times, accounting for phylogenetic and phenotypic uncertainty, using the R packages ape (version `r packageVersion('ape')`)[@paradis2004ape] and nlme (version `r packageVersion('nlme')`)[@pinheiro2014r]. In these analyses we used a Brownian Motion based covariance matrix for traits.

For regressions at the species and genus level, we reshuffled and matched records at each iteration to account for variation across records for the same taxon. For regressions at the family level we recalculated the average ovariole number per insect family, downsampling the representation for each family by half. No posterior distribution was made available with the published family level phylogeny[@rainford2014phylogenetic].

```{r PGLS-setup}
# build body size dataset
# set factor for resampling the body size dataframe, here no downsampling used
downsample_factor <- 1.0
fam_count_threshold <- 1
source("analysis/combine_body_egg_datasets.R")
source("analysis/pgls_functions.R") # get functions for plotting PGLS results
```

>>>>

---
title: Repeated loss of variation in insect ovary morphology highlights the role of developmental constraint in life-history evolution - Supplementary Methods
output:
    bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: pdflatex
    #bookdown::html_document2: default
toc: true
bibliography: manuscript_supplement.bib
csl: nature.csl
link-citations: yes
header-includes:
 \usepackage{float}
 \usepackage{array}
 \usepackage{xcolor,colortbl}
 \floatplacement{figure}{H}
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(echo=FALSE, include = FALSE, warning=FALSE, message=FALSE, cache=FALSE)
```

```{r preliminaries}
library(BAMMtools)
library(phytools)
library(coda)
library(OUwie)
library(corHMM)
library(grid)
library(gridExtra)
library(gtable)
library(geiger)
library(arbutus)
library(dplyr)
library(ggtree)
library(ape)
library(ggplot2)
library(RColorBrewer)
theme_set(theme_classic())
options(tinytex.clean = FALSE)
```

@@

```{r load_data}
# This script reads in and formats the ovariole number dataset
source("analysis/get_ovariole_number_data.R")
```

Samuel H. Church^1^\*, Bruno A. S. De Medeiros^1,2^, Seth Donoughe^1,3^, Nicole L. Márquez Reyes^4^, Cassandra G. Extavour^1,5^\*

@@

Using this approach, we gathered `r nrow(raw_ON)` records for ovariole number from 460 publications. A full list of publications is provided in the supplementary file 'ovariole_number_bibliography.pdf'. We matched the scientific names to additional taxonomic information using the software TaxReformer[@TaxReformer] and found additional taxonomic data for `r nrow(matched)` of the `r nrow(raw_ON)` records. We verified that TaxReformer had found a valid match by comparing the originally recorded taxonomic order to the order populated by online databases, and removed `r nrow(raw_ON %>% filter(problem == "taxonomy"))` taxonomic records for which these values did not match. For all subsequent analyses, we also excluded observations made in non-reproductive individuals from eusocial species (workers), as well as two observations which represented significant outliers and could not be validated using additional sources or figures[@su2015testicular;@hernandez2017biology].

@@

Average adult body length per insect family[@rainford2016phylogenetic] was matched to the average ovariole number for the corresponding families (Fig. \ref{fig:PGLS-family-length}). The Rainford et al, 2016[@rainford2016phylogenetic] dataset contains a small number of average adult body lengths at the order level (e.g. Strepsiptera), which were matched to their equivalent group in the ovariole number dataset. To test the effect of uncertainty in the estimated average ovariole number on our results, the dataset for each family was downsampled by half at each iteration of the regression analysis.

@@

```{r PGLS-setup}
source("analysis/pgls_functions.R")

# combine datasets, using all data and randomly shuffled database entries
downsample_factor <- 1
fam_count_threshold <- 1

set.seed(12345)  # set seed for cobmination
source("analysis/combine_body_egg_datasets.R")

# use the maximum clade credibility tree
tree <- genus_mcc_tree
```

```{r PGLS-egg-ON}
### PGLS egg volume vs ON, species level pgls
pgls_results_egg_ON_name <- run_all_taxa_pgls(egg_ON_name %>% 
    select(genus,logvol,logon,group) %>% 
    rename(rank = genus, trait1 = logon, trait2 = logvol) ,
    tree)

egg_ON_name_plot <- egg_ON_name %>% 
    select(genus,vol,on,group) %>% 
    rename(rank = genus, trait1 = on, trait2 = vol)

pdf("manuscript_figures/panels/FigEgg_ON_name.pdf",height=5,width=7.5,useDingbats = F)
g1 <- ggplot(egg_ON_name_plot,aes(x = trait1, y = trait2, color = group)) +
    geom_point() +
    scale_color_manual(values = mrk) +
    scale_x_log10(breaks=c(1,10,100,1000,10000)) +
    scale_y_log10(label=scientific_10,breaks=c(0.00001,0.0001,0.001,0.01,0.1,1,10,100,100)) +
    theme(legend.position = "none") +
    xlab('ovariole number') +
    ylab(bquote('egg volume '(~mm^3)))
print(g1)
invisible(dev.off())
```

```{r PGLS-egg-ON-genus, fig.cap="**Egg volume vs ovariole number, matching records at the genus level.**", fig.width=7.5, fig.height=5, include=TRUE}
### PGLS on vs egg volume by insect genus
egg_ON_genus_pgls <- egg_ON_genus %>% 
    select(genus,logvol,logon,group) %>% 
    rename(rank = genus, trait1 = logon, trait2 = logvol)

pgls_results_egg_ON_genus <- run_all_taxa_pgls(egg_ON_genus_pgls,tree)

run_plotting_pgls(egg_ON_genus_pgls,tree,"egg_ON_genus") +
    xlab(bquote('ovariole number,'~log[10])) +
    ylab(bquote('egg volume,'~mm^3~','~log[10]))
```

```{r PGLS-egg-ON-weight-genus}
### Gilbert egg vs ON, ind body weight genus pgls
gilbert_genus_pgls <- egg_ON_gilbert_genus %>% 
    select(genus,logvol,logon,logweight,group) %>% 
    mutate(rank = genus, trait1 = logon, trait2= logvol, indep = logweight)

pgls_results_gilbert_genus <- run_all_taxa_resid_pgls(gilbert_genus_pgls,tree)

### Gilbert egg volume vs ON, independent body weight, genus level pgls
pdf("manuscript_figures/panels/FigEgg_gilbert_weight_residuals.pdf",height=2.5,width=2.5,useDingbats = F)
run_plotting_resid_pgls(gilbert_genus_pgls,tree,"gilbert_genus") + 
    xlab("residual of ovariole number") + 
    ylab("residual of egg volume")
invisible(dev.off())
```

```{r PGLS-Gilbert-name, fig.cap="**Egg volume vs ovariole number, phylogenetic residuals to dry adult body weight, matching records at the species level.**", fig.width=3, fig.height=3, include=TRUE}
### Gilbert egg vs ON, ind body weight name pgls
egg_ON_gilbert_name_pgls <- egg_ON_gilbert_name %>% 
    select(genus,logvol,logon,logweight,group) %>% 
    mutate(rank = genus, trait1 = logon, trait2= logvol, indep = logweight)

pgls_results_gilbert_name <- run_all_taxa_resid_pgls(egg_ON_gilbert_name_pgls,tree)

### Gilbert body weight, species level pgls
run_plotting_resid_pgls(egg_ON_gilbert_name_pgls,tree,"gilbert_name") + 
    xlab("residual of ovariole number") + 
    ylab("residual of egg volume")
```

```{r PGLS-family-length, fig.cap="**Family average egg volume vs ovariole number, phylogenetic residuals to adult body length.**",  fig.width=3, fig.height=3, include=TRUE}
### PGLS egg vs ON, independent body size
body_egg_ON_family_pgls <- body_egg_ON_full %>% 
    select(rank,logvol,logon,logbody,group) %>% 
    rename(trait1 = logon, trait2= logvol, indep = logbody)

pgls_results_body_egg_ON_family <- run_all_taxa_resid_pgls(body_egg_ON_family_pgls,fam_tree)

run_plotting_resid_pgls(body_egg_ON_family_pgls,fam_tree,"body_egg_ON_family") + 
    xlab("residual of average ovariole number") +
    ylab("residual of average egg volume")
```

\textcolor{red}{We ran each Phylogenetic Generalized Least Squares regression over the Maximum Clade Credibilty (MCC) tree. For each regression, we also repeated each PGLS analysis 1000 times,} accounting for phylogenetic and phenotypic uncertainty, using the R packages ape (version `r packageVersion('ape')`)[@paradis2004ape] and nlme (version `r packageVersion('nlme')`)[@pinheiro2014r]. In these analyses we used a Brownian Motion based covariance matrix for traits.

For regressions at the species and genus level, we reshuffled and matched records at each iteration to account for variation across records for the same taxon. For regressions at the family level we recalculated the average ovariole number per insect family, downsampling the representation for each family by half. No posterior distribution was available with the previously published family level phylogeny[@rainford2014phylogenetic].

\textcolor{red}{For regressions of egg size and ovariole number when accounting for adult body size, we compared the results of our regression analyses to distributions estimated using simulated data under alternative hypotheses. We fit a Brownian motion model to the phylogenetic residuals of egg size and body size (R package geiger}[@harmon2007geiger]\textcolor{red}{, version `r packageVersion('geiger')`), and then used the parameters of this fitted model to simulate new datasets (R package phylolm}[@tung2014linear]\textcolor{red}{, version `r packageVersion('phylolm')`). We performed this resimulation using the datasets of egg size and body length at the family level, and egg size and body weight at the genus level. We simulated 1000 datasets each under two hypotheses: no correlation (slope=0) and a strong negative correlation (slope=-1:1). We performed the regressions as described above and compared the distribution of p-values and slopes to values from regressions on observed data.}

```{r simulated, fig.cap="**\\textcolor{red}{Using simulated data to test alternative hypotheses of evolutionary relationships}** Top row, distributions of p-values over 1000 replicate regressions, dashed black line indicates threshold of 0.01. Bottom row, distribution of estimated slopes between egg size and ovariole number, dashed lines indicate 95% interval of simulated distributions. Left, comparing egg size and ovariole number, accounting for body length at the family level. Right, comparing egg size and ovariole number, accounting for body weight at the genus level. Red=observed values, gray=simulated with no correlation, black=simulated with a -1:1 correlation. n=1000 regressions.", fig.width=7.5, fig.height=5, include=TRUE}
get_tstat <- function(sum,n) {
    tstat <- as.numeric(unname(sum[[n]]$coefficients[[2]]))
    return(tstat)
}

get_pval <- function(sum,n) {
    pval <- as.numeric(unname(sum[[n]]$tTable[[8]]))
    return(pval)
}

load("analysis/pgls_results_Fall2020/run_pgls_simulated_allometry_body_egg_ON_family_-1_image.RData")
sim_corr_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=1)
sim_corr_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=1)
load("analysis/pgls_results_Fall2020/run_pgls_simulated_allometry_body_egg_ON_family_0_image.RData")
sim_nocorr_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=1)
sim_nocorr_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=1)
load("analysis/pgls_results_Fall2020/run_posterior_pgls_egg_size_ovariole_number.RData")
obs_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=3)
obs_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=3)

body_egg_ON_family_tstat_dist <- rbind(tibble(type = "simulated",
    relationship = "-1:1 correlation",
    tstat=sim_corr_tstat_dist),
tibble(type = "simulated",
    relationship = "no correlation",
    tstat=sim_nocorr_tstat_dist),
tibble(type = "observed",
    relationship = "observed",
    tstat=obs_tstat_dist))

body_egg_ON_family_pval_dist <- rbind(tibble(type = "simulated",
    relationship = "-1:1 correlation",
    pval=sim_corr_pval_dist),
tibble(type = "simulated",
    relationship = "no correlation",
    pval=sim_nocorr_pval_dist),
tibble(type = "observed",
    relationship = "observed",
    pval=obs_pval_dist))

load("analysis/pgls_results_Fall2020/run_pgls_simulated_allometry_gilbert_genus_-1_image.RData")
sim_corr_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=1)
sim_corr_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=1)
load("analysis/pgls_results_Fall2020/run_pgls_simulated_allometry_gilbert_genus_0_image.RData")
sim_nocorr_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=1)
sim_nocorr_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=1)
load("analysis/pgls_results_Fall2020/run_posterior_pgls_egg_size_ovariole_number.RData")
obs_tstat_dist <- sapply(allometry_distribution_raw,get_tstat,n=5)
obs_pval_dist <- sapply(allometry_distribution_raw,get_pval,n=5)

gilbert_genus_pval_dist <- rbind(tibble(type = "simulated",
    relationship = "-1:1 correlation",
    pval=sim_corr_pval_dist),
tibble(type = "simulated",
    relationship = "no correlation",
    pval=sim_nocorr_pval_dist),
tibble(type = "observed",
    relationship = "observed",
    pval=obs_pval_dist))

gilbert_genus_tstat_dist <- rbind(tibble(type = "simulated",
    relationship = "-1:1 correlation",
    tstat=sim_corr_tstat_dist),
tibble(type = "simulated",
    relationship = "no correlation",
    tstat=sim_nocorr_tstat_dist),
tibble(type = "observed",
    relationship = "observed",
    tstat=obs_tstat_dist))

t1 <- ggplot(body_egg_ON_family_tstat_dist,aes(x=tstat,y=relationship,color=relationship)) + geom_jitter(size=0.2)  + 
geom_vline(xintercept=quantile(body_egg_ON_family_tstat_dist %>% filter(relationship == "-1:1 correlation") %>% pull(tstat),0.05),colour="black",linetype="longdash") + 
geom_vline(xintercept=quantile(body_egg_ON_family_tstat_dist %>% filter(relationship == "-1:1 correlation") %>% pull(tstat),0.95),colour="black",linetype="longdash") + 
geom_vline(xintercept=quantile(body_egg_ON_family_tstat_dist %>% filter(relationship == "no correlation") %>% pull(tstat),0.05),colour="dark gray",linetype="longdash") + 
geom_vline(xintercept=quantile(body_egg_ON_family_tstat_dist %>% filter(relationship == "no correlation") %>% pull(tstat),0.95),colour="dark gray",linetype="longdash") + 
#ggtitle("size vs number, body length, family")  + 
xlab("test statistic, slope") +
#ylab("type") + 
xlim(c(-2,2)) +
scale_color_manual(values = c("black","dark gray","red")) +
theme(legend.position="none",
    axis.title.y=element_blank())

t2 <- ggplot(gilbert_genus_tstat_dist,aes(x=tstat,y=relationship,color=relationship)) + geom_jitter(size=0.2)  + 
geom_vline(xintercept=quantile(gilbert_genus_tstat_dist %>% filter(relationship == "-1:1 correlation") %>% pull(tstat),0.05),colour="black",linetype="longdash") + 
geom_vline(xintercept=quantile(gilbert_genus_tstat_dist %>% filter(relationship == "-1:1 correlation") %>% pull(tstat),0.95),colour="black",linetype="longdash") + 
geom_vline(xintercept=quantile(gilbert_genus_tstat_dist %>% filter(relationship == "no correlation") %>% pull(tstat),0.05),colour="dark gray",linetype="longdash") + 
geom_vline(xintercept=quantile(gilbert_genus_tstat_dist %>% filter(relationship == "no correlation") %>% pull(tstat),0.95),colour="dark gray",linetype="longdash") + 
#ggtitle("size vs number, body weight, genus") + 
xlab("test statistic, slope") +
#ylab("type") + 
xlim(c(-2,2)) +
scale_color_manual(values = c("black","dark gray","red")) +
theme(legend.position="none",
    axis.title.y=element_blank(),
axis.text.y=element_blank(),
axis.ticks.y=element_blank())

p1 <- ggplot(body_egg_ON_family_pval_dist,aes(x=pval,y=relationship,color=relationship)) + geom_jitter(size=0.2)  + 
geom_vline(xintercept=0.01,colour="black",linetype="dashed") + 
#geom_vline(xintercept=0.05,colour="black",linetype="longdash") + 
ggtitle("egg size vs ov. number, body length")  + 
xlab("p-value") +
#ylab("type") + 
xlim(c(0,1)) +
scale_color_manual(values = c("black","dark gray","red")) +
theme(legend.position="none",
    axis.title.y=element_blank())

p2 <- ggplot(gilbert_genus_pval_dist,aes(x=pval,y=relationship,color=relationship)) + geom_jitter(size=0.2)  + 
geom_vline(xintercept=0.01,colour="black",linetype="dashed") + 
#geom_vline(xintercept=0.05,colour="black",linetype="longdash") + 
ggtitle("egg size vs ov. number, body weight") + 
xlab("p-value") +
#ylab("type") + 
xlim(c(0,1)) +
scale_color_manual(values = c("black","dark gray","red")) +
theme(legend.position="none",
    axis.title.y=element_blank(),
axis.text.y=element_blank(),
axis.ticks.y=element_blank())

grid.arrange(p1,p2,t1,t2,ncol=2,widths=5:4)
```

```{r PGLS-sublineage}
### Apoidea ON, egg, indep body pgls
DIP_body_egg_ON_name_pgls <- DIP_body_egg_ON_name %>% 
    select(name,logvol,logon,logbody,group) %>% 
    rename(rank = name, trait1 = logon, trait2= logvol, indep = logbody)

pgls_results_DIP <- run_all_taxa_resid_pgls(DIP_body_egg_ON_name_pgls,dros_tree)

pdf("manuscript_figures/panels/FigEgg_Drosophila_residuals.pdf",height=2.5,width=2.5,useDingbats = F)
run_plotting_resid_pgls(DIP_body_egg_ON_name_pgls,dros_tree,"DIP") + 
    xlab("residual of ovariole number") +
    ylab("residual of egg volume")
invisible(dev.off())

### Orthoptera ON, egg, indep body pgls
ORT_body_egg_ON_genus_pgls <- ORT_body_egg_ON_genus %>% 
    select(genus,logvol,logon,logbody,group) %>% 
    rename(rank = genus, trait1 = logon, trait2= logvol, indep = logbody)

pgls_results_ORT <- run_all_taxa_resid_pgls(ORT_body_egg_ON_genus_pgls,tree)

pdf("manuscript_figures/panels/FigEgg_Orthoptera_residuals.pdf",height=2.5,width=2.5,useDingbats = F)
run_plotting_resid_pgls(ORT_body_egg_ON_genus_pgls,tree,"ORT") + 
    xlab("residual of ovariole number") +
    ylab("residual of egg volume")
invisible(dev.off())
```

```{r PGLS-Coleoptera-Hymenoptera, fig.cap="**Additional lineage-specific comparisons of egg volume vs ovariole number, phylogenetic residuals to body size, matching records at the genus level.** Weevils (Curculionoidea, left) were measured using elytra length and wasps (Hymenoptera, right) were measured using mesosoma width.", fig.width=6, fig.height=3, include=TRUE}
### Curculionoidea ON, egg, indep body pgls
COL1_body_egg_ON_genus_pgls <- COL1_body_egg_ON_genus %>% 
    select(genus,logvol,logon,logbody,group) %>% 
    rename(rank = genus, trait1 = logon, trait2 = logvol, indep = logbody)

pgls_results_COL1 <- run_all_taxa_resid_pgls(COL1_body_egg_ON_genus_pgls,tree)

theme_update(plot.title = element_text(hjust = 0.5))
col1 <- run_plotting_resid_pgls(COL1_body_egg_ON_genus_pgls,tree,"COL1") + 
    xlab("residual of ovariole number") +
    ylab("residual of egg volume") + 
    ggtitle("Coleoptera: Curculionoidea")

### Apoidea ON, egg, indep body pgls
HYM_body_egg_ON_genus_pgls <- HYM_body_egg_ON_genus %>% 
    select(genus,logvol,logon,logbody,group) %>% 
    rename(rank = genus, trait1 = logon, trait2 = logvol, indep = logbody)

pgls_results_HYM <- run_all_taxa_resid_pgls(HYM_body_egg_ON_genus_pgls,tree)

hym1 <- run_plotting_resid_pgls(HYM_body_egg_ON_genus_pgls,tree,"HYM") + xlab("residual of ovariole number") +
ylab(element_blank()) + ggtitle("Hymenoptera")

```{r PGLS-body-plot, fig.cap="**Adult body size vs ovariole number.** Adult body weight compared at the genus level (left), and average adult body length compared at the family level (right).", fig.width=6, fig.height=3, include=TRUE}
### body length vs ON, family level
ON_body_family_pgls <- body_egg_ON_full %>% 
    select(rank,logon,logbody,group) %>% 
    rename(trait1 = logon, trait2 = logbody)

pgls_results_body_ON_family <- run_all_taxa_pgls(ON_body_family_pgls,fam_tree)

family_body <- run_plotting_pgls(ON_body_family_pgls,fam_tree,"body_egg_ON_family") + 
    xlab(bquote('average ovariole number,'~log[10])) +
    ylab(bquote('average adult body length (mm),'~log[10]))

### Gilbert body weight vs ON, species level
pgls_results_body_gilbert_name <- run_all_taxa_pgls(egg_ON_gilbert_name %>% 
    select(genus,logon,logweight,group) %>% 
    mutate(rank = genus, trait1 = logon, trait2 = logweight) ,
    tree)

### Gilbert body weight, genus level
ON_body_gilbert_genus_pgls <- egg_ON_gilbert_genus %>% 
    select(genus,logon,logweight,group) %>% 
    mutate(rank = genus, trait1 = logon, trait2 = logweight)

pgls_results_body_gilbert_genus <- run_all_taxa_pgls(ON_body_gilbert_genus_pgls,tree)

gilbert_body <- run_plotting_pgls(ON_body_gilbert_genus_pgls,tree,"body_gilbert") + xlab(bquote('ovariole number,'~log[10])) +
ylab(bquote('adult dry body weight (g),'~log[10]))

### Curculionoidea body vs ON
pgls_results_body_COL1 <- run_all_taxa_pgls(COL1_body_egg_ON_genus %>% 
    select(genus,logon,logbody,group) %>% 
    rename(rank = genus, trait1 = logon, trait2 = logbody) ,
    tree)

### Orthoptera body vs ON
pgls_results_body_ORT <- run_all_taxa_pgls(ORT_body_egg_ON_genus %>% 
    select(genus,logon,logbody,group) %>% 
    rename(rank = genus, trait1 = logon, trait2 = logbody) ,
    tree)

### Apoidea body vs ON
pgls_results_body_HYM <- run_all_taxa_pgls(HYM_body_egg_ON_genus %>% 
    select(genus,logon,logbody,group) %>% 
    rename(rank = genus, trait1 = logon, trait2 = logbody) ,
    tree)

### Drosophilidae body vs ON
pgls_results_body_DIP <- run_all_taxa_pgls (DIP_body_egg_ON_name %>% 
    select(name,logon,logbody,group) %>% 
    rename(rank = name, trait1 = logon, trait2 = logbody) ,
    dros_tree)

```{r PGLS-fecundity-name, fig.cap="**Lifetime fecundity vs ovariole number, matching records at the species level.**", fig.width=3, fig.height=3, include=TRUE}
### Gilbert fecundity vs ON name
fecundity_name_pgls <-  ON_gilbert_name %>% 
    select(genus,logfecundity,logon,group) %>% 
    mutate(rank = genus, trait1 = logon, trait2= logfecundity)

pgls_results_fecundity_name <- run_all_taxa_pgls(fecundity_name_pgls,tree)

run_plotting_pgls(fecundity_name_pgls,tree,"fecundity_name") + 
    xlab(bquote("ovariole number,"~log[10])) +
    ylab(bquote("lifetime fecundity,"~log[10]))
```

```{r PGLS-fecundity-genus}
### Gilbert fecundity vs ON genus
fecundity_genus_pgls <-  ON_gilbert_genus %>% 
    select(genus,fecundity,on,group) %>% 
    mutate(rank = genus, trait1 = on, trait2= fecundity)

pgls_results_fecundity_genus <- run_all_taxa_pgls(fecundity_genus_pgls,tree)

pdf("manuscript_figures/panels/FigFecundity.pdf",height=4,width=4,useDingbats = F)
g1 <- ggplot(fecundity_genus_pgls,aes(x = trait1, y = trait2, color = group)) +
    geom_point() +
    scale_color_manual(values = mrk) +
    scale_x_log10() +
    scale_y_log10() +
    theme(legend.position = "none") +
    xlab('ovariole number') +
    ylab('lifetime fecundity')
print(g1)
invisible(dev.off())
```

```{r PGLS-posterior-egg-ON, include=TRUE}
# retrieve results from over the posterior distribution
get_pgls_table <- function(dist) {
    table <- round(data.frame(
    slope_min = min(sapply(dist,function(x) {x$coefficients[[2]]})),
    slope_max = max(sapply(dist,function(x) {x$coefficients[[2]]})),
    pval_sig = sum(sapply(dist,function(x) {x$tTable[[8]] < 0.01})),
    taxa = min(sapply(dist,function(x) {x$dims$N}))),
    4)
    return(table)
}

@@

### PGLS TABLE EGG SIZE OVARIOLE NUMBER
load("analysis/pgls_results_Fall2020/run_posterior_pgls_egg_size_ovariole_number.RData")

egg_ON_posterior_results <- list(allometry_egg_ON_name,
    allometry_egg_ON_genus,
    allometry_gilbert_name,
    allometry_gilbert_genus,
    allometry_body_egg_ON_family)

egg_ON_pgls_results <- list(pgls_results_egg_ON_name,
    pgls_results_egg_ON_genus,
    pgls_results_gilbert_name,
    pgls_results_gilbert_genus,
    pgls_results_body_egg_ON_family)

egg_ON_pgls_analysis_names <- c("ovariole number vs egg volume",
    "ovariole number vs egg volume",
    "ovariole number vs egg volume, residuals to body weight",
    "ovariole number vs egg volume, residuals to body weight",
    "ovariole number vs egg volume, residuals to body length")

egg_ON_taxonomic_levels <- c("species","genus","species","genus","family")

egg_ON_table <- as.data.frame(bind_rows(lapply(egg_ON_posterior_results,get_pgls_table))) %>% mutate(analysis = egg_ON_pgls_analysis_names,
        "taxonomic level" = egg_ON_taxonomic_levels,
        "slope" = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
        "MCC p-value" = gsub("0.000","<0.001",sprintf("%.3f",sapply(egg_ON_pgls_results,function(x){x$tTable[[8]]}))),
        "num. sig. / 1000" = pval_sig,
        "taxa" = taxa)
knitr::kable(egg_ON_table %>% select(c("analysis","taxonomic level","slope","MCC p-value","num. sig. / 1000","taxa")),
        "latex",
        caption="Results of PGLS analysis of ovariole number and egg size across a posterior distribution", booktabs = T, linesep = "") %>% 
    kableExtra::kable_styling(latex_options = "striped") %>%
    kableExtra::column_spec(1, width = "16em") %>%
    kableExtra::kable_styling(latex_options = "hold_position")  %>%
    kableExtra::kable_styling(font_size = 8) 
```

```{r PGLS-posterior-clade, include=TRUE}
### PGLS TABLE SUBCLADE
load("analysis/pgls_results_Fall2020/run_posterior_pgls_subclade.RData")

clade_posterior_results <- list(allometry_DIP,
    allometry_ORT,
    allometry_COL1,
    allometry_HYM)

clade_pgls_results <- list(pgls_results_DIP,
    pgls_results_ORT,
    pgls_results_COL1,
    pgls_results_HYM)

clade_pgls_analysis_names <- c("Drosophilidae ovariole number vs egg volume, residuals to thorax length",
    "Orthoptera ovariole number vs egg volume, residuals to body length",
    "Curculionoidea ovariole number vs egg volume, residual to elytra length",
    "Hymenoptera ovariole number vs egg volume, residuals to mesosoma width")

clade_taxonomic_levels <- c("species","genus","genus","genus")

clade_table <- as.data.frame(bind_rows(lapply(clade_posterior_results,get_pgls_table))) %>% mutate(analysis = clade_pgls_analysis_names,
        "taxonomic level" = clade_taxonomic_levels,
        "slope" = paste(sprintf("%.3f",slope_min),sprintf("%.3f",slope_max),sep=sp),
        "MCC p-value" = gsub("0.000","<0.001",sprintf("%.3f",sapply(clade_pgls_results,function(x){x$tTable[[8]]}))),
        "num. sig. / 1000" = pval_sig,
        "taxa" = taxa)
knitr::kable(clade_table %>% select(c("analysis","taxonomic level","slope","MCC p-value","num. sig. / 1000","taxa")),
        "latex",
        caption="Results of PGLS analysis of ovariole number and egg size across a posterior distribution", booktabs = T, linesep = "") %>% 
    kableExtra::kable_styling(latex_options = "striped") %>%
    kableExtra::column_spec(1, width = "16em") %>%
    kableExtra::kable_styling(latex_options = "hold_position")  %>%
    kableExtra::kable_styling(font_size = 8) 
```
